Study of Higgs plus single top production using events with a same sign dimuon at the Large Hadron Collider

### 2019-11-24 tesis_hiram/Chapter2/chapter2.tex
correccion

The LHC is designed to accelerate particles to high energies and generate collisions. There are several
experiments installed along the LHC ring as shown in figure \ref{lhc}. One of them is ATLAS, located on Point 1 between the two injection lines, which is a general purpose detector. At point 5 is another general purpose detector: the compact muon solenoid (CMS). A particle detector optimized for heavy ion physics, ALICE, is located at Point 2 and LHCb, a detector designed for B physics, is located at Point 7\cite{cern1,cern2}.

@@

@@

\begin{figure}[ht]
\centering
\includegraphics[width=16cm,height=10cm]{Chapter1/cern.jpg}
\caption[Large hadron collider from CERN]{CERN accelerator complex \cite{cern1}} \label{lhc}
\end{figure}

%\begin{table}[ht]
%\centering
%	\caption[Accelerator operation energies]{Accelerator operation energies\cite{cern3,cern1}}
%	\begin{tabular}{|c|c|}
%		\hline
%		Accelerator & Energy \\
%		\hline
%		Linac 2 & 50 MeV \\
%		\hline
%		PS Booster & 1.4 GeV \\
%		\hline
%		Proton Scyncroton (PS) & 25 GeV\\
%		\hline
%		SPS & 450 GeV\\
%		\hline
%		LHC & 6.5 TeV\\
%		\hline
	%\end{tabular}
%\end{table}
\pagebreak
Inside the detectors, colliding protons generate an amount of particles such as pions and kaons, which are the most common, created by the jet particles that were produced after the collision. Besides the particles created by the quarks of the proton, gluons are also radiated and create new particles and photons are radiated too. 
During the collision time, the main detectors of LHC capture the energy 
and save the data in supercomputers. Those high end computers starts to measure the signals in order to assign a type of events.

The main objective of the LHC is to study the nature of electroweak symmetry breaking for the Higgs mechanism of the SM. 
Alternatives to the Standard Model that involves different symmetries are also put to test. With the exploration of theses theories, people hope for a discovery that guides them towards a unified theory, so the importance to accelerate particles to higher energy scales.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{Chapter2/lum6.png}
\caption[LHC schedule, including future plans for increasing the center of mass energy and luminosity. The periods labeled LS are the maintenance periods]{LHC schedule, including future plans for increase the center of mass energy and luminosity. The periods labeled LS are the maintenance periods\cite{cern3}.}
\label{lhc-lumi}
\end{figure}
Figure \ref{lhc-lumi} shows the Large Hadron Collider forecast for the increase of luminosity for the next years. Red dots represent the expected peak luminosity and blue line shows the expected luminosity. In the year 2018, the maximum integrated luminosity is around 150 $fb^{-1}$\cite{cern3}. From 2019 to 2021 the second phase of maintenance will take place, where engineers increase the performance of the accelerator, give maintenance to the system and introduce new components to the complex. After that period, the LHC will start a new collision period at even higher energies in order to explore new particle phenomena.

\section{The Compact Muon Solenoid}	
The Compact Muon Solenoid (CMS) is a detector with multiple uses in the LHC and part of the main experiments at CERN. It is located underground in the France- Switzerland border, in the city of Cessy, France.
This detector was designed in the early 1990s, based on the mass limit of the Higgs boson, and put on operation in 2008. It has a big solenoid that generates a great magnetic field of 4 teslas with the objective of separating particles after a particle collision.
 The detector is 21 meters long, 15 meters wide, 15 meters high, it has a diameter of 5.9 m with a weight of 12000 ton. The reason for such a strong magnetic field is to obtain a better momentum resolution.
 The magnetic flux is returned via a 1.5 m thick iron yoke instrumented with four stations of muon
 chambers. %Characteristics of the solenoid are shown in table \ref{tab:my_label}.

\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/cms2.png}
	\caption[View of Compact muon solenoid (CMS)]{View of Compact muon solenoid (CMS)\cite{cms-manual}}\label{cms}
\end{figure}
There are many layers of detectors to register different kinds of energy signals. But since muons go through materials very easily, the complex is buried 100 meters underground.

The CMS experiment is composed of several detector layers that allow identify and save different types of energy signals and it is saved in a powerful supercomputers that separate and classify the data according to certains variables. The general parts of the CMS are the Silicon Tracker, an Electromagnetic Calorimeter, a Hadron Calorimeter, the solenoid (superconducting magnet) and the Muon Detector. All these parts are illustrated in figure \ref{cms}.

\begin{table}[ht]
	\caption[Characteristic of the CMS superconducting solenoid]{Characteristic of the CMS superconducting solenoid\cite{cms-manual}}
	\centering
	\begin{tabular}{|c|c|}
		\hline
		Field strength & 4 T \\
		Inner Bore & 5.9 m \\
		Length & 12.9 m\\
		Number of Turns & 2168\\
		Current & 19.5 kA\\
		Stored energy & 2.7 GJ\\
		\hline
	\end{tabular}
	\label{tab:my_label}
\end{table}

%The trigger or event selection process must reduce the approximately 1 billion interactions per second to more less than 100 events per second for storage and subsequent analysis.\\

@@

The Strip Detector has almost 15 400 modules, mounted on
carbon-fiber structures and housed inside a temperature controlled outer support tube. The operating temperature must be around -$\ang{20}$C. 
The total area of the pixel detector is around 1 $m^2$ and the silicon strips is 200 m$^2$. 
The inner tracker comprises 66 million pixels and 9.6
million silicon strips \cite{cms-manual}.
A full coverage of the Tracker is shown in figure \ref{pixel2}

\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.7]{Chapter2/pixel2.png}
	\caption[The Tracker layout distributed in terms of $\eta$]{The Tracker layout distributed in terms of $\eta$\cite{cms-manual}.}
	\label{pixel2}
\end{figure}
The particle collision generates a lot of different particles that pass first through the inner tracker, interacting with the sensor layers and registering the particle path.
By reconstructing the path of the particles, it is possible to generate a track, which allows to measure the momentum of the particle by calculating its curvature. The tracker is used to reconstruct the path of charged particles (e.g. electrons, muons, pions). The most important objects of study are the momentum and the vertex (origin of the track). The efficiency of the tracker is estimated by using samples of muons and pions with $p_t$ that is the transverse momentum (component of momentum transverse to beam line) of 1,10 and 100 GeV. 
For muons the efficiency of the tracker is shown in figure \ref{efi}
\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/tracker.png}
	\caption[Total reconstruction efficiency for muons]{Total reconstruction efficiency for muons\cite{cms-manual} }\label{efi}
\end{figure}

According to figure \ref{efi}, the efficiency of the tracker is around 98 $\%$ for $|\eta|<2$. The resolution of the tracker for muons with $p_t$=100 GeV is around 1-2 $\%$.
Of course, this is a first stage of the measurement of particle momentum because the other detectors are used to increase and improve the measurement and reconstruction of the particle path.

@@

The endcaps cover the range 1.479 < $|\eta|$ < 3.0. The distance between the interaction point and the endcap is 3.144 m. The endcap has identically shaped crystals grouped in mechanical units of 5×5 crystals. The crystal cross section for the rear face is 30 $\times$ 30 mm$^2$, and for the face is 28.62 $\times$ 28.62 mm$^2$ with length of 220 mm. The ECAL is illustrated in figure \ref{ecal}.

@@

As its name says, the ECAL is mainly used for the electron detection, but also detects, photons, and neutral pions. The charged particles reach the crystals that provokes the photo diodes detect energy.
The resolution of the ECAL is estimated by using a test beam and registering the energy signals. The resolution for electrons with energies of 120 GeV is 0.5$\%$.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/hcal.png}
	\caption[View of the CMS HCAL showing the different layers and tower regions. In the endcap, the towers are defined with two longitudinal segments as shown by the colors]{View of the CMS HCAL showing the different layers and tower regions. In the endcap, the towers are defined with two longitudinal segments as shown by the colors\cite{cms-manual}}\label{hcal}
\end{figure}
The outer barrel hadron calorimeter (HO) consists of layers of scintillator with thickness of 10 mm, located outside of the
magnet coil that cover the region -1.26 < $|\eta|$ < 1.26.
There is also a forward calorimeter (HF) located at 11.2 m from the interaction point, that cover the region of 2.9 < $|\eta|$ < 5. They are made of
steel absorber and embedded radiation hard quartz fibers, which collect
Cherenkov light\cite{cms-manual}.% that is when a charged particle travels through matter faster than light can, resulting in a blue light emmision of the material. 
%The length of the fibers are 1.65 m and 1.43 m, which are alterned with a 5 mm of separation.

@@

\subsection{Muon Detector}
%The muon detector system %has 3 primary functions: muon
%triggering, identification, and momentum measurement\cite{cms7}. 
%The process utilized in the CMS muon chamber systems is gas ionization. Those gas-ionization particle detectors are arranged in a cylindrical barrel region and planar endcaps, in order to meet the geometry of the solenoid. The barrel detector is built with 4 layers of 250 chambers inside the magnet return yoke.\cite{cms-manual}
The Muon Detector has three main subdetectors: the drift tube chambers (DT), the cathode strips chambers (CSC) and the resistive plate chamber (RPC). 
The DT's, CSC's and RPC's are shown in figure \ref{dt}.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=11cm,height=7cm]{Chapter2/csc}
	\caption[Cross section of a quadrant of the CMS with axis parallel to the beam (z axis) horizontally and the radius of the detector in terms of the $\eta$]{Cross section of a quadrant of the CMS with axis parallel to the beam (z axis) horizontally and the radius of the detector in terms of the $\eta$\cite{cms-manual}.}
	\label{dt}
\end{figure}

@@

%The produced muons are measured in three zones: in the inner tracker, after the coil, and in
%the return flux. Measurement the momentum of muons in the muon detector is determined by the muon bending angle at the exit of the 4T coil with the interaction point as the origin of the muon\cite{cms7}\cite{cms-manual}.
%The resolution of the muom detector is based on simulations taking account with the geometry and a test beam. 
The main purpose of the Muon System Detector is to identify muon tracks. 
The efficiency of the muon system is estimated by using single muon samples simulated with $p_t$=10, 50, 100, 500 and 1000 GeV. The results are shown in the figure \ref{mu-efi}.
\begin{figure}
	\centering
	\includegraphics[width=0.6\linewidth]{Chapter2/mu-efi}
	\caption[Muon reconstruction efficiency in terms of $\eta$ and $p_t$]{Muon reconstruction efficiency in terms of $\eta$ and $p_t$\cite{cms-manual}}
	\label{mu-efi}
\end{figure}
The graph shows an efficiency of approximately 98$\%$ in the the fully instrumented regions. In the intersection regions, there are drops in efficiency due to the separation of the chambers. % There is a reduction of the efficiency to 6$\%$, which corresponds to the $\eta$ position for the separation.

>>>>

The LHC is designed to accelerate particles to high energies and generate collisions. There are several
experiments installed along the LHC ring as shown in Figure \ref{lhc}. One of them is ATLAS, located on Point 1 between the two injection lines, which is a general purpose detector. At point 5 is another general purpose detector: the compact muon solenoid (CMS). A particle detector optimized for heavy ion physics, ALICE, is located at Point 2 and LHCb, a detector designed for B physics, is located at Point 7\cite{cern1,cern2}.

@@

\afterpage{
\begin{figure}
	\centering
	\includegraphics[width=16cm,height=15cm]{Chapter1/cern.jpg}
	\caption[CERN accelerator complex]{CERN accelerator complex \cite{cern1}} \label{lhc}
\end{figure}
\clearpage}
%\begin{table}[ht]
%\centering
%	\caption[Accelerator operation energies]{Accelerator operation energies\cite{cern3,cern1}}
%	\begin{tabular}{|c|c|}
%		\hline
%		Accelerator & Energy \\
%		\hline
%		Linac 2 & 50 MeV \\
%		\hline
%		PS Booster & 1.4 GeV \\
%		\hline
%		Proton Scyncroton (PS) & 25 GeV\\
%		\hline
%		SPS & 450 GeV\\
%		\hline
%		LHC & 6.5 TeV\\
%		\hline
	%\end{tabular}
%\end{table}

Inside the detectors, colliding protons generate an amount of particles such as pions and kaons, which are the most common, created by the jet particles that were produced after the collision. Besides the particles created by the quarks of the proton, gluons are also radiated and create new particles and photons are radiated too. 
Within a short time, the main detectors of LHC capture 
and save the data using high performance computing.

The main objective of the LHC is to study the nature of electroweak symmetry breaking for the Higgs mechanism of the SM. 
Alternatives to the Standard Model that involves different symmetries are also put to test. With the exploration of theses theories, people hope for a discovery that guides them towards a unified theory, so the importance to accelerate particles to higher energy scales and high intensity experiments.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{Chapter2/lum6.png}
\caption[LHC schedule]{LHC schedule, including future plans for increase the center of mass energy and luminosity. The periods labeled LS are the maintenance periods\cite{cern3}.}
\label{lhc-lumi}
\end{figure}
Figure \ref{lhc-lumi} shows the Large Hadron Collider forecast for the increase of luminosity for the next years. Red dots represent the expected peak luminosity and blue line shows the expected luminosity. In the year 2018, the maximum integrated luminosity is around 150 $fb^{-1}$\cite{cern3}. From 2019 to 2021 the second phase of maintenance will take place, where engineers increase the performance of the accelerator, give maintenance to the system and introduce new components to the complex. After that period, the LHC will start a new collision period at even higher energies in order to explore new particle phenomena.

@@

\section{The Compact Muon Solenoid}	
The Compact Muon Solenoid (CMS) is a detector with multiple uses in the LHC and part of the main experiments at CERN. It is located underground in the France- Switzerland border, in the city of Cessy, France.
This detector was designed in the early 1990s, based on the mass limit of the Higgs boson, and put on operation in 2008. It has a big solenoid that generates a great magnetic field of 4 teslas with the objective of separating particles after a particle collision.
 The detector is 21 meters long, 15 meters wide, 15 meters high, it has a diameter of 5.9 m with a weight of 12000 ton. The reason for such a strong magnetic field is to obtain a better momentum resolution. More characteristics of the solenoid are shown in Table \ref{tab:my_label}.  
 \begin{table}[ht]
 	\caption[Characteristic of the CMS superconducting solenoid]{Characteristic of the CMS superconducting solenoid\cite{cms-manual}}
 	\centering
 	\begin{tabular}{|c|c|}
 		\hline
 		Field strength & 4 T \\
 		Inner Bore & 5.9 m \\
 		Length & 12.9 m\\
 		Number of Turns & 2168\\
 		Current & 19.5 kA\\
 		Stored energy & 2.7 GJ\\
 		\hline
 	\end{tabular}
 	\label{tab:my_label}
 \end{table}

 %The magnetic flux is returned via a 1.5 m thick iron yoke instrumented with four stations of muon
 %chambers. 
\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/cms2.png}
	\caption[View of Compact muon solenoid (CMS)]{View of Compact muon solenoid (CMS)\cite{cms-manual}}\label{cms}
\end{figure}
%There are many layers of detectors to register different kinds of energy signals. %But since muons go through materials very easily, the complex is buried 100 meters underground. 
The CMS experiment is composed of several detector layers that allow identify and save different types of energy signals and it is saved in a powerful supercomputers that separate and classify the data according to certains variables. The general parts of the CMS are the Silicon Tracker, an Electromagnetic Calorimeter, a Hadron Calorimeter, the solenoid (superconducting magnet) and the Muon Detector. All these parts are illustrated in Figure \ref{cms}.
%The trigger or event selection process must reduce the approximately 1 billion interactions per second to more less than 100 events per second for storage and subsequent analysis.\\

@@

The Strip Detector has almost 15 400 modules, mounted on
carbon-fiber structures and housed inside a temperature controlled outer support tube. The operating temperature must be around -$\ang{20}$C. 
The total area of the pixel detector is around 1 $m^2$ and the silicon strips is 200 m$^2$. 
The inner tracker comprises 66 million pixels and 9.6
million silicon strips \cite{cms-manual}.
A full coverage of the Tracker is shown in Figure \ref{pixel2}

\begin{figure}[ht]
	\centering
	\includegraphics[scale=0.7]{Chapter2/pixel2.png}
	\caption[The Tracker layout distributed in terms of $\eta$]{The Tracker layout distributed in terms of $\eta$\cite{cms-manual}.}
	\label{pixel2}
\end{figure}
The particle collision generates a lot of different particles that pass first through the inner tracker, interacting with the sensor layers and registering the particle path.
By reconstructing the path of the particles, it is possible to generate a track, which allows to measure the momentum of the particle by calculating its curvature. The tracker is used to reconstruct the path of charged particles (e.g. electrons, muons, pions). The most important objects of study are the momentum and the vertex (origin of the track). The efficiency of the tracker is estimated by using samples of muons and pions with $p_t$ that is the transverse momentum (component of momentum transverse to beam line) of 1,10 and 100 GeV. 
For muons the efficiency of the tracker is shown in Figure \ref{efi}
\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/tracker.png}
	\caption[Total reconstruction efficiency for muons]{Total reconstruction efficiency for muons\cite{cms-manual} }\label{efi}
\end{figure}

According to Figure \ref{efi}, the efficiency of the tracker is around 98 $\%$ for $|\eta|<2$. The resolution of the tracker for muons with $p_t$=100 GeV is around 1-2 $\%$.
Of course, this is a first stage of the measurement of particle momentum because the other detectors are used to increase and improve the measurement and reconstruction of the particle path.

@@

The endcaps cover the range 1.479 < $|\eta|$ < 3.0. The distance between the interaction point and the endcap is 3.144 m. The endcap has identically shaped crystals grouped in mechanical units of 5×5 crystals. The crystal cross section for the rear face is 30 $\times$ 30 mm$^2$, and for the face is 28.62 $\times$ 28.62 mm$^2$ with length of 220 mm. The ECAL is illustrated in Figure \ref{ecal}.

@@

As its name says, the ECAL is mainly used for the electron detection, but also detects, photons, and neutral pions. The charged particles reach the crystals, create a shower of secondary electrons and photons. The light from the secondary particles is detected by  photo diodes.
The resolution of the ECAL is estimated by using a test beam and registering the energy signals. The resolution for electrons with energies of 120 GeV is 0.5$\%$.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=10cm,height=6cm]{Chapter2/hcal.png}
	\caption[View of the HCAL ]{View of the CMS HCAL showing the different layers and tower regions. In the endcap, the towers are defined with two longitudinal segments as shown by the colors\cite{cms-manual}}\label{hcal}
\end{figure}
The outer barrel hadron calorimeter (HO) consists of layers of scintillator with thickness of 10 mm, located outside of the
magnet coil that cover the region -1.26 < $|\eta|$ < 1.26.
There is also a forward calorimeter (HF) located at 11.2 m from the interaction point, that cover the region of 2.9 < $|\eta|$ < 5. They are made of
steel absorber and embedded radiation hard quartz fibers, which collect
Cherenkov light\cite{cms-manual}.% that is when a charged particle travels through matter faster than light can, resulting in a blue light emmision of the material. 
%The length of the fibers are 1.65 m and 1.43 m, which are alterned with a 5 mm of separation.

@@

\subsection{Muon Detector}
%The muon detector system %has 3 primary functions: muon
%triggering, identification, and momentum measurement\cite{cms7}. 
%The process utilized in the CMS muon chamber systems is gas ionization. Those gas-ionization particle detectors are arranged in a cylindrical barrel region and planar endcaps, in order to meet the geometry of the solenoid. The barrel detector is built with 4 layers of 250 chambers inside the magnet return yoke.\cite{cms-manual}
The Muon Detector has three main subdetectors: the drift tube chambers (DT), the cathode strips chambers (CSC) and the resistive plate chamber (RPC). 
The DT's, CSC's and RPC's are shown in Figure \ref{dt}.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=11cm,height=7cm]{Chapter2/csc}
	\caption[CMS Muon detector]{Cross section of a quadrant of the CMS with axis parallel to the beam (z axis) horizontally and the radius of the detector in terms of the $\eta$, illustrating the muon detector and its components \cite{cms-manual}.}
	\label{dt}
\end{figure}

@@

%The produced muons are measured in three zones: in the inner tracker, after the coil, and in
%the return flux. Measurement the momentum of muons in the muon detector is determined by the muon bending angle at the exit of the 4T coil with the interaction point as the origin of the muon\cite{cms7}\cite{cms-manual}.
%The resolution of the muom detector is based on simulations taking account with the geometry and a test beam. 
The main purpose of the Muon System Detector is to identify muon tracks. 
The efficiency of the muon system is estimated by using single muon samples simulated with $p_t$=10, 50, 100, 500 and 1000 GeV. The results are shown in figure \ref{mu-efi}.
\begin{figure}
	\centering
	\includegraphics[width=0.6\linewidth]{Chapter2/mu-efi}
	\caption[Muon reconstruction efficiency in terms of $\eta$ and $p_t$]{Muon reconstruction efficiency in terms of $\eta$ and $p_t$\cite{cms-manual}}
	\label{mu-efi}
\end{figure}
The graph shows an efficiency of approximately 98$\%$ in the the fully instrumented regions. In the intersection regions, there are drops in efficiency due to the separation of the chambers. % There is a reduction of the efficiency to 6$\%$, which corresponds to the $\eta$ position for the separation.
